\section{Fluid Description}
	This section aims to provide an overview of the derivation of the fluid equations
	from a Kinetic Theory perspective. First the Vlasov equation is introduced,
	then it is explained how the fluid equations can be obtained by taking different
	order moments of the Vlasov equation. This can help understand the limitations
	of the fluid model of plasma. Lastly a few different approximations
	to make the fluid equations closable.

\subsection{Kinetic Theory}
	Let \(\mathcal{F}_s\) be the exact phase-space density of a particle species,
	it contains all the positions, velocities for all the particles at
	all times. By integrating over all velocities and multiplying with the charge
	for all species we obtain the charge density, \(\rho_c\).

	\[\rho_c = \sum_s e_s \int \mathcal{F}_s(\vb{r},\vb{v},t)\dd[3]{\vb{v}}\]

	Likewise we find he current density, \(\vb{j}\).

	\[\vb{j} = \sum_s e_s \int \vb{v}\mathcal{F}_s(\vb{r},\vb{v},t)\dd[3]{\vb{v}}\]

	Then its seems we can derive all plasma interaction from considering
	the conservation of the phase-space density, coupled with Maxwells equations.
	The phase-space conservation is given by what is known as Vlasovs equation \cref{eq:vlasov}
	\citep{pecseli_waves_2012}).

	\begin{align}
		\pdv{\mathcal{F}}{t} + \vb{v}\cdot\nabla\mathcal{F}_s + \frac{e_s}{m_s}\left( \vb{E} + v\cross\vb{B} \right) \cdot \nabla_v\mathcal{F}_s = 0 \label{eq:vlasov}
	\end{align}

	where \(\nabla_v\) is the velocity grad-operator.
	Unfortunately this expression, combined with Maxwells equations, is only solvable
	for special simple geometries.

\subsection{Velocity Moments}
	To investigate plasma as a fluid we have to make certain fluid approximations.
	The plasma is then characterized by local parameters describing particle
	density, kinetic temperature, flow velocity and so on. These parameters refer
	to a small volume of plasma, in contrast with the discussions earlier \cref{sec:single_particle}.
	The time evolution is then governed by the fluid equation, but unfortunately the resulting
	equations are generally less tractable than the usual hydrodynamical
	equations.

	The usual quantity known as the momemt is given by mass times velocity,
	here we will introduce a more general form of moment where the usual moment
	is the first order moment. This will help understand how the fluid
	equations result from averaging over different moments of the general
	transport equation. The zeroth, first and second order moment is respectively
	given by

	\begin{subequations}
		\begin{equation}
			\Phi^0(\vb{v}) = m
		\end{equation}
		\begin{equation}
			\Phi^1(\vb{v}) = m\vb{v}
		\end{equation}
		\begin{equation}
			\Phi^2(\vb{v}) = m\vb{v}\vb{v}
		\end{equation}
	\end{subequations}

	By integrating the moment functions and the distribution function
	\(\mathcal{F}\), over the velocity space we can retrieve different quantities.

	Integrating the zeroth order moment gives the density, if we divide by the
	mass.

	\begin{equation}
			n = \frac{1}{m}\int m \mathcal{F} \dd\vb{v}
	\end{equation}

	Integrating over the first order moment gives the momentum, if we divide
	by density.

	\begin{equation}
			m\vb{v} = \int m \vb{v} \mathcal{F} \dd\vb{v}
	\end{equation}

	We can in fact find the mean of any order moment by integrating the
	distribution function over \(\mathcal{F}\).

	\begin{equation}
		\left< \Phi^n(\vb{v}) \right> = \frac{1}{n} \int \Phi^n \mathcal{F} \dd\vb{v}
	\end{equation}

\subsection{Transport Equation}
	By multiplying the a momentum function, \( \Phi \), with the Vlasov equation, \cref{eq:vlasov},
	we obtain the general momentum transport equation. (Need Citation for this, sent mail and asked.)

	\begin{equation}
		\pdv{n\left< \Phi^n(\vb{v}) \right>}{t} + \nabla \cdot \left( \left< \Phi^n(\vb{v})\vb{v} \right> \right)
		= \frac{n}{m} \left< \vb{F}_L \cdot \nabla_v \Phi^n(\vb{v}) \right>
	\end{equation}

	This then becomes a conservation equation for the average macroscopic quantity
	\(\left< \Phi \right>\). By using this equation and inserting in the zeroth, first and second order
	moment we obtain the fluid equations, \cref{eq:continuity,eq:momentum,eq:pressure}.
	We refer you to \citetitle{fitzpatrick_plasma_2014} to see a derivation of the fluid
	equations.


	% To simplify matters we can average over an ensemble to obtain the
	% average distribution function \(\mathcal{\bar{F}}_s\),
	% \[\mathcal{\bar{F}}_s \equiv \left< \mathcal{F}_s\right>_{ensemble}\].
	%
	% The averaged distribution is then used in \cref{eq:vlasov} to make it more tractable.
	% The electric, \(\vb{E}\), and magnetic fields, \(\vb{B}\), is statistically
	% dependent of the distribution function and this leads to correlations when
	% using the averaged distribution. We will not go into those here, so see
	% \textit{\citetitle{fitzpatrick_plasma_2014}}\citep{fitzpatrick_plasma_2014}
	% for a treatment of it.


\subsection{Fluid Equations}
	\label{sec:fluid}
	The generalized fluid equations is given in \cref{eq:continuity,eq:momentum,eq:pressure}, where the three equations
	describe the conservation of mass, momentum and energy respectively. The collision term is negleted.
	We refer you to the earlier mentioned \textit{\citetitle{fitzpatrick_plasma_2014}}, although some
	notation differ, for the rather involved derivation to obtain them.

	\begin{subequations}
		\begin{equation}
			\left( \pdv{t} + \vb{u}_s\cdot\nabla \right)n_s + n_s\nabla\cdot \vb{u}_s =0
			\label{eq:continuity}
		\end{equation}
		\begin{equation}
			m_sn_s\left( \pdv{t} + \vb{u}_s \cdot \nabla \right)\vb{u}_s = - \nabla p_s - \nabla \cdot \pi  + n_s\vb{f}_s
			\label{eq:momentum}
		\end{equation}
		\begin{equation}
			\left( \pdv{t} + \vb{u_s} \cdot \nabla \right)p_s =
			- \frac{5}{3} p_s\nabla \cdot \vb{u}_s -
			\frac{2}{3}\pi_s : \nabla{\vb{u_s}}
			- \frac{2}{3}\nabla \cdot \vb{q}_s
			\label{eq:pressure}
		\end{equation}
	\end{subequations}

	% (NOTE TO SELF: NEED TO DEFINE PROPERLY ALL TERMS)
	% To simplify matters we can average over an ensemble to obtain the
	% average distribution function \(\mathcal{\bar{F}}_s\),
	% \[\mathcal{\bar{F}}_s \equiv \left< \mathcal{F}_s\right>_{ensemble}\].
	%
	% The averaged distribution is then used in \cref{eq:vlasov} to make it more tractable.
	% The electric, \(\vb{E}\), and magnetic fields, \(\vb{B}\), is statistically
	% dependent of the distribution function and this leads to correlations when
	% using the averaged distribution. We will not go into those here, so see
	% \textit{\citetitle{fitzpatrick_plasma_2014}}\citep{fitzpatrick_plasma_2014}
	% for a treatment of it.

	The first equation equation, \cref{eq:continuity}, is the continuity equation, it states that the total mass
	in a volume	should be preserved. \(\vb{u}_s\) is the flow velocity and \(n_s\) is the number density, i.e.
	number of particles in a volume. The divergence terms signifies change due to the compressability of the fluid
	and can in many cases be set to \(0\). The total derivative, i.e. \(\left(\pdv{t} + \vb{u}_s\cdot\nabla \right)\) accounts for
	change in density in a volume taking into account substance exiting and entering.
	The momentum equation \cref{eq:momentum} shows that the fluid momentum change, left hand side,
	is due to pressure gradients, \(\nabla p_s\), visceous forces, \(\nabla \cdot \pi \) and external forces, \(n_s\vb{f}_s\),
	per unit volume. (NOTE: Should I add in precise mathy definition of viscosity/pressure/heat?)
	Lastly we have the energy equation, in its pressure form, which shows that changes to thermal
	energy, \(p = nkT\), is caused by compression, \(p_s\nabla \cdot \vb{u}_s\), visceous effects \(\pi_s : \nabla{\vb{u_s}}\)
	and heat transport	\(\frac{2}{3}\nabla \cdot \vb{q}_s\).
	The fluid equations is in general not closeable and adding higher order moments
	always introduces more unknowns. Due to this one generally uses different closing
	schemes, some of which is described in the next section, to make them tractable.

\subsection{Plasma States}
	Plasma can be classified according to which	approximations and simplifications
	that are valid for them. Here we will go through a few of them.

	\subsubsection{Local Thermodynamic Equilibrium}
	A plasma is said to be in a \textit{local thermodynamic equalibrium} (LTE)
	if the phase-space distribution is locally maxwellian. This means the variations
	in temperature is slow enough that we can consider it to flow no heat. We can also ignore
	the viscosity due to there being little local variations to the momentum flow.

	\begin{align}
		\mathcal{F}_m = \frac{n}{(2\pi )^{3/2}v_t^3} \exp{-\frac{(v-u)^2}{2v_t^2}}
	\end{align}

	Since the viscosity tensor, \(\vb{\pi}\), and the heat flux tensor, \(\vb{q}\)
	contains odd integrals over the distribution, \textit{\citetitle[see][]{fitzpatrick_plasma_2014}},
	they dissappear.

	\subsubsection{Cold Plasma}
	In what we consider a cold plasma the pressure, \(p\), and viscosity, \(\pi\), is set to zero.
	This can be useful if the velocities of interest far exceed the thermal velocities.
	(ADD MORE)


	\subsubsection{Isothermal}
	An isothermal plasma is one where we assume an infinite heat conductivity,
	this means the temperatures is constant in all space and time. This can be used to
	describe macroscopic plasma.
	(ADD MORE)
