\section{Langmuir Oscillations}
	\label{sec:langmuir}

	Plasma oscillations, also called Langmuir oscillations, is the basic resulting
	oscillation	that happens as a plasma tries to reach a stable equilibrium, due to
	a small perturbation of its density.
	We will use this to show how the fluid equations can be closed for a simple system
	using assumptions. This is also very suited to test simulations, as we do later
	in \cref{sec:langmuir_verification}.

	Here we will consider an one specie plasma fluid consisting of electrons under local thermal equilibrium, LTE.
	The electron density, \(n_0\) and pressure, \(p_0\), is initally homogenous.
 	The fluid has a vanishing flow, \(\va{u}_0 = 0\), and no initial potential gradient \(\phi_0 = 0\).
	The ele
 	See \citet{pecseli_waves_2012} for a another discussion.

	A small perturbation of the electron density will cause the electric field
	to try to restore the equilibrium. When the electrons reach the equilibrium
	position they will have a kinetic energy and will overshoot.This will cause
	a new perturbation away from the equilibrium.

	Under the LTE conditions the fluid equations simplify to

	\begin{subequations}
		\begin{equation}
		\pdv{n_e}{t} + \nabla \cdot (n_e \va{u}_e) = 0
		\end{equation}
		\begin{equation}
		m_en_e\left( \pdv{}{t} + \va{u}_e \cdot \nabla \right)\va{u}_e = e n_e \nabla \phi - \nabla p_e
		\end{equation}
		\begin{equation}
		\left( \pdv{}{t} + \va{u}_e \cdot \nabla \right) p_e + \frac{5}{3} p_e\nabla \cdot \va{u}_e = 0
		\end{equation}
	\end{subequations}
	Since this set of equations have more unknowns than equations so we need additional information
	to close the set. Here we can use the Poisson equation to close it.
	\begin{equation}
	\epsilon_0 \nabla^2 \phi = e\left( n_e - n_0 \right)
	\end{equation}

	Now we let a small perturbation, denoted by a tilde, happen to the equilibrium.
	Since we are free to choose an inertial reference frame, we select one co-moving
	with the plasma so the inital fluid velocity is \(\va{u}_0=0\). We also select
	the reference potential so the initial potential, \(\phi_0\),  is \(0\).


	\begin{equation*}
	\text{Perturbation} \rightarrow
	\begin{cases}
	  n_e = n_0 + \tilde{n}_e\\
	  p_e = p_0 + \tilde{p}_e\\
	  \va{u}_e = \tilde{\va{u}}_e\\
	  \phi = \tilde {\phi}
	\end{cases}
	\end{equation*}

	Since the pertubation is small, we can say that any part that contains
	second order terms of perturbation of a quantity will be much smaller than the value
	of the quantity, \(q \gg\tilde q \tilde q\). So even though we may miss some processes
	by doing this, we can drop the second order perturbation terms.
	This process is called linearization.

	Inserting the perturbation and linearizing the equations we get:

	\begin{subequations}
	  \begin{equation}
	    \pdv{\tilde{n}_e}{t} + \nabla \cdot (n_0 \tilde{\va{u}}_e) = 0 \label{eq:langmuir_continuity}
	  \end{equation}
	  \begin{equation}
	    m_e \pdv{\tilde{\va{u}}_e}{t}  = e  \nabla \tilde{ \phi} - \frac{\nabla \tilde{p}_e}{n_0}
	  \end{equation}
	  \begin{equation}
	     \pdv{\tilde{p}}{t} +\frac{5}{3}p_0 \nabla \cdot \tilde{\va{u}}_e = 0 \label{eq:langmuir_energy}
	  \end{equation}
	  \begin{equation}
	    \epsilon_0 \nabla^2 \tilde{\phi} = e\tilde{n}_e
	  \end{equation}
	\end{subequations}

	Then we combine the continuity and energy equations, \cref{eq:langmuir_continuity} and \cref{eq:langmuir_energy}.

	\begin{align}
	\pdv{}{t}\left( \frac{\tilde{p}_e}{p_0} + \frac{5}{3} \frac{\tilde{n}_e}{n_0} \right) &= 0
	\end{align}

	The perturbed pressure and density are proportional, \(\nabla \tilde{p}_e = \left(5p_0/3n_0\right)\nabla \tilde{n}_e\).
	Assuming plane wave solutions along the x-axis, the differential operators become \(\nabla \rightarrow ik\)
 	and \(\pdv{}{t} \rightarrow -i\omega\), we can solve for the dispersion relation.

	\begin{align}
	\epsilon(\omega, k) = 1 + \frac{5}{3} \lambda_{se}^2k ^2 -  \frac{\omega^2}{\omega_{pe}^2}
	\end{align}

	Here the electron debye length, \(\lambda_{D}\), and the plasma frequency, \(\omega_{pe}\), have been inserted.

% \section{Langmuir Oscillations}
% 	REDO, You have already done this in a neater way before!!!!!!!!!!!!
%
% 	In a we consider a homogenous and isotropic plasma, in stable equilibrium,
% 	and let the electrons be pushed, causing a slight perturbation of the equilibrium.
% 	The slightly uneven distribution will cause an electric field pushing against
% 	the perturbation and try to restore the equilibrium. When the electrons reach
% 	the equilibrium position they have a kinetic energy and will overshoot
%  	causing an equal opposite perturbation. Then it repeats and we have a simple
% 	oscillation of the electron density.
%
% 	Certain assumptions are necessary to derive the oscillation mathematically.
% 	First the plasma needs to be in a homogenous and isotropic equilibrium state
%  	so the spatial and temporal derivatives is zero. The magnetic field strength
% 	also needs to be small enough to be safely ignored.	We then consider movements
% 	on a timescale so that the inertial effects of the electrons are important,
% 	while the ions are considered stationary.
%
% 	We start from the electron fluid motion equation,
% 	\begin{equation}
% 	mn_e \left( \pdv{}{t} + \vec{u} \cdot \nabla \right)\vec{u} = -en_e\vec{E} - \nabla p_e
% 	\end{equation}
% 	and we consider a small perturbation to the equilibrium state so the quantities becomes:
% 	\begin{equation}
% 	\vec{u} \approx \vec{u}_0 + \vec{\tilde{u}}; \qquad{}
% 	\vec{E} \approx \tilde{\vec{E}}; \qquad{}
% 	n \approx n_0 + \tilde{n};\qquad{}
% 	p \approx p_0 + \tilde {p}
% 	\end{equation}
% 	Here the subscript for the electron is dropped, the subscript \(0\) is
%  	the equilibrium state and the tilde is the perturbation. Then we apply
% 	linearization to the equation, so that all the second order terms of the
% 	perturbation goes away.
% 	\begin{equation}
% 	mn_0\pdv{}{t} \tilde{\vec{u}} = -en_0\vec{\tilde{E}} - \nabla \tilde{p_e}
% 	\end{equation}
% 	For simplicities sake the perturbation is a plane wave in the x-direction, \(exp[i(kx - i\omega)]\),
% 	as well as what we will program for verification use. Then the differential
% 	operators become \(\nabla \rightarrow ik\) and \(\pdv{}{t} \rightarrow -i\omega\),
% 	using the relation \(\tilde{p} = {3T\tilde{n}}\), see cite{Goldston INtro to plasma 1995}.
% 	Then the x-component of the electron motion equation becomes:
% 	\begin{equation}
% 	i\omega mn_0 \tilde{u} = e n_0 \tilde{E} + i 3 kt\tilde{n}
% 	\end{equation}
%
% 	Using the same procedure the electron continuity equation,
% 	\(\pdv{n}{t} + \nabla \cdot (\vec{u} n) = 0 \), becomes
% 	\begin{equation}
% 		-i\omega \tilde{n} = ikn_0 \tilde{u}
% 	\end{equation}
%
% 	Next we look poisson's equation, \(\epsilon_0 \nabla \cdot \vec{E} = e(n_i-n_e) \), which using the same procedures,
% 	as well as letting the ion density cancel the equilibrium electron density,	ends up as
% 	\begin{equation}
% 		ik\epsilon_0\tilde{E} = -\tilde{n}
% 	\end{equation}
