\section{Single Particle Motion}
	To better understand the collective motion of plasma it is useful to consider
	the motions of the single particles that the plasma consists of. By at first
	treating only one particle we can ignore the electromagnetic influence from
	the other particles which greatly simplifies the situation. The Lorentz
	force, \(\vb{F}_L = q\left(\vb{E} + \vb{v} \cross \vb{B} \right)\), (Check which terms need an introduction) is often
	the dominant force in a plasma due to it's relative strength
 	compared to gravity and other forces.
 	\begin{align}
		\vb{F} = q\left(\vb{E} + \vb{v}\cross\vb{B}\right) \label{eq:lorentz}
	\end{align}
	To simplify matters we will only consider particles in static electromagnetic field,
	as that is often a valid approximation on the time and spatial scales particle gyration
	operates on. There several other types of drift, and motions, that appear in various configurations,
	that we will not go through here, i.e. gradient-B drift, curvature drift, polarization drift and magnetic mirror
	to mention a few.


	\subsection{Gyration}
		\label{sec:gyration}
		Let us consider a situation with a single particle in a static and isotropic external
 		magnetic field \(\vb{B}\), with initial velocity \(\vb{v}\).
		The equation of motion for the particle will is then

		\begin{align}
			m\pdv{\vb{v}}{t} &= q\vb{v}\cross\vb{B} \label{eq:simple_gyration}
			\intertext{The velocity parallel to the magnetic field is unaffected by the field. So we will
			seperate the velocity into a parallel and perpendicular velocity with respect to the magnetic field,
			\(\vb{v}=\vb{v}_\parallel + \vb{v}_\perp\). The parallel motion is fully described by
			\(\pdv{\vb{v}_\parallel}{t} = \frac{q}{m}\vb{E_\parallel}\).
			We remove the parallel part and perform temporal derivative on the equation of motion to get }
			\pdv[2]{\vb{v}_\perp}{t} &= \frac{q}{m} \pdv{\vb{v}_\perp}{t}\cross \vb{B}
			\intertext{Then we insert \cref{eq:simple_gyration} into the equation and use the vector relation
			\( a\cross b\cross c = b(a\cdot b) - c(a\cdot b) \).}
			\pdv[2]{\vb{\omega}_\perp}{t} + \left(\frac{qB}{m}\right)^2\vb{\omega}_\perp &= 0
		\end{align}

		This differential equation corresponds to a gyration around the magnetic field lines
		with the gyration frequency, \(\omega_c = \frac{qB}{m}\), as the frequency.
		In the last equation we also changed the term describing the rotational motion
		to \(\vb{\omega}_\perp\), which is often used in plasma literature to
		signify gyration.

	\subsection{E-cross-B Drift}
	The E-cross-B drift happens when a particle is moving within static and isotrop
	electric and magnetic fields. The equation of motion, neglecting all forces
	except the electromagnetic, is then

	\begin{align}
		m\pdv{v}{t} &= q(\vb{E} + \vb{v}\cross\vb{B}) \label{eq:EcrossB}
	\end{align}

	As in the previous section we decompose the velocity into a parallel and perpendicular
 	velocity. We also assume there exists a constant drift \(\vb{v}_D\), somewhat unfounded at this stage,
	so that we can seperate the perpendicular motion into the drift and the gyration.
	The velocity is then given by \(\va{v} = \va{v}_\parallel + \va{\omega}_\perp + \va{v}_D\),
	and we insert this into the equation of motion.

	\begin{align}
		m\pdv{}{t}\left( \va{v}_\parallel + \va{\omega}_\perp + \va{v}_D\right) &=
		q \left( \va{E} +   \left( \va{v}_\parallel + \va{\omega}_\perp +
		\va{v}_D \right)  \cross \va{B}_0 \right)
	\end{align}


	As before the parallel velocity is unaffected by the magnetic field we subtract
	it from the equation. From \cref{sec:gyration} we know that the gyration
	part is given by

	\begin{align}
		m\pdv{\vb{\omega}_\perp}{t} &= q\vb{\vb{\omega}_\perp}\cross\vb{B}
	\end{align}

	After we have subtracted these two parts we are left with

	\begin{align}
		\pdv{\va{v}_D}{t} &= \frac{q}{m} \left( \va{E}_\perp + \va{v}_D \cross \va{B} \right)
	\end{align}

	Then we use the previous assumption that the drift velocity is constant,
	cross the equation with \(\vb{B}\) and perform some algebra to arrive at

	\begin{align}
		\vb{v}_D &= \frac{\vb{E}\cross\vb{B}}{B^2}
	\end{align}

	As we can see the E-cross-B drift is independent of particle charge and mass,
	which means both the ions and electrons will be drifting at the same velocity
	subject to the drift which will be perpendicular to electric and magnetic fields.
